---
title: "acesinabcd"
author: "FB"
date: "2024-03-11"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Youth-deferred ACEs proxy scores for ABCD data

Running this code will create an ACEs proxy score for Baseline, Yr1, Yr2, and Yr3 with preference given to the youth (vs parent/caregiver) report when available.
Need to run the setup file (ELAfromABCD_setup.Rmd) first to create the necessary RDS and CSV files.
IMPORTANT: You will need to run the ENTIRE code to get correct scores for Years 1-3, as they build on each other.

```{r load-packages}
library(dplyr)
library(tidyr)
```

```{r load-data}

#read RDS file
ela_in_abcd=readRDS("ela_in_abcd.rds")
mh_py_le_bs=read.csv("mh_py_le_bs.csv")

#create baseline dataset
ela_in_abcd_bs = ela_in_abcd[ela_in_abcd$eventname == "baseline_year_1_arm_1", ]

#add in life events "baseline" data
#remove the variable label from main data set and re-add without eventname so the calculated baseline appears inline with other baseline data
ela_in_abcd_bs = ela_in_abcd_bs[c("src_subject_id","eventname","ksads_ptsd_raw_761_p","ksads_ptsd_raw_762_p","ksads_ptsd_raw_763_p",  "ksads_ptsd_raw_764_p","ksads_ptsd_raw_765_p","ksads_ptsd_raw_767_p","ksads_ptsd_raw_768_p","ksads_ptsd_raw_769_p","ksads_ptsd_raw_766_p","ksads_ptsd_raw_756_p","ksads_ptsd_raw_757_p","ksads_ptsd_raw_758_p","ksads_ptsd_raw_759_p","ksads_ptsd_raw_760_p","fes_youth_q5","crpbi_y_ss_parent_ace","demo_fam_exp1_v2","demo_fam_exp6_v2","demo_fam_exp7_v2","asr_scr_totprob_t","asr_q06_p","asr_q06_p_ace","asr_scr_totprob_t_ace","famhx_ss_momdad_ma_p","famhx_ss_momdad_alc_p","famhx_ss_momdad_dg_p","famhx_ss_momdad_dprs_p","famhx_ss_momdad_hspd_p","famhx_ss_momdad_scd_p","famhx_ss_momdad_vs_p")]
ela_in_abcd_bs = left_join(ela_in_abcd_bs, mh_py_le_bs, by=c("src_subject_id")) 

#create separate dfs for remaining time points
ela_in_abcd_yr1 = ela_in_abcd[ela_in_abcd$eventname == "1_year_follow_up_y_arm_1", ]
ela_in_abcd_yr2 = ela_in_abcd[ela_in_abcd$eventname == "2_year_follow_up_y_arm_1", ]
ela_in_abcd_yr3 = ela_in_abcd[ela_in_abcd$eventname == "3_year_follow_up_y_arm_1", ]

#creating function to sum items in each domain where it will sum across the row as long as at least one variable is not NA
sum_score <- function(x){

  if (all(is.na(x))) {

    suma <- NA
  }  
  else {    
    suma <- sum(x, na.rm = T)
  }

  return(suma)
}

```


```{r baseline-yd}
#abuse_phy
#only parent report available
#KSADS	Shot, stabbed, or beaten brutally by a non-family member	ksads_ptsd_raw_761_p
#KSADS	Shot, stabbed, or beaten brutally by a grown up in the home	ksads_ptsd_raw_762_p
#KSADS	Beaten to the point of having bruises by a grown up in the home	ksads_ptsd_raw_763_p
#KSADS	A non-family member threatened to kill your child	ksads_ptsd_raw_764_p
#KSADS	A family member threatened to kill your child	ksads_ptsd_raw_765_p

#create dataframe with phys abuse items
abuse_phy_bs_data = c("ksads_ptsd_raw_761_p" , "ksads_ptsd_raw_762_p", "ksads_ptsd_raw_763_p", "ksads_ptsd_raw_764_p", "ksads_ptsd_raw_765_p")

#sum the items as long as they're not all NA
ela_in_abcd_bs$abuse_phy_sum = apply(ela_in_abcd_bs[abuse_phy_bs_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_bs$abuse_phy_yd_bs = ifelse(ela_in_abcd_bs$abuse_phy_sum == 0, 0, 1)

#abuse_sex
#only parent report available
#KSADS A grown up in the home touched your child in their privates, had your child touch their privates, or did other sexual things to your childP ksads_ptsd_raw_767_p
#KSADS An adult outside your family touched your child in their privates, had your child touch their privates or did other sexual things to your childP ksads_ptsd_raw_768_p
#KSADS A peer forced your child to do something sexuallyP ksads_ptsd_raw_769_p

#create dataframe with sexual abuse items
abuse_sex_bs_data = c("ksads_ptsd_raw_767_p" , "ksads_ptsd_raw_768_p", "ksads_ptsd_raw_769_p")

#sum the items as long as they're not all NA
ela_in_abcd_bs$abuse_sex_sum = apply(ela_in_abcd_bs[abuse_sex_bs_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_bs$abuse_sex_yd_bs = ifelse(ela_in_abcd_bs$abuse_sex_sum == 0, 0, 1)

#abuse_emo
#youth report used for main ACEs variable as well
#Family Environment Family members often criticize each other. fes_youth_q5
#utilizing youth report only to ensure youth exposure (e.g., not parents criticizing without kid present)

#yes/no item; no sum or recode necessary
ela_in_abcd_bs$abuse_emo_yd_bs = ela_in_abcd_bs$fes_youth_q5

#neglect_phy
#only parent report available
#Demographics Needed food but couldn't afford to buy it or couldn't afford to go out to get it?P demo_fam_exp1_v2_l/demo_fam_exp1_v2
#Demographics Had someone who needed to see a doctor or go to the hospital but didn't go because you could not afford it?P demo_fam_exp6_v2/demo_fam_exp6_v2_l
#Demographics Had someone who needed a dentist but couldn't go because you could not afford it?P demo_fam_exp7_v2/demo_fam_exp7_v2_l

#create dataframe with physical neglect items
neglect_phy_bs_data = c("demo_fam_exp1_v2", "demo_fam_exp6_v2", "demo_fam_exp7_v2")

#sum the items as long as they're not all NA
ela_in_abcd_bs$neglect_phy_sum = apply(ela_in_abcd_bs[neglect_phy_bs_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_bs$neglect_phy_yd_bs = ifelse(ela_in_abcd_bs$neglect_phy_sum == 0, 0, 1)


#neglect_emo
#youth report used for main ACEs proxy variable as well
#CRPBI parent *CRPBI Mean Parent (primary caregiver)Y crpbi_y_ss_parent_ace

#all recoding done previously
ela_in_abcd_bs$neglect_emo_yd_bs = ela_in_abcd_bs$crpbi_y_ss_parent_ace


#divorce
#Life Events Parents separated or divorced? ple_separ_y
ela_in_abcd_bs$divorce_yd_bs = ela_in_abcd_bs$ple_separ_y_bs


#violence in home (violence)
#only parent report available
#KSADS Witness the grownups in the home push, shove or hit one anotherP 	ksads_ptsd_raw_766_p
ela_in_abcd_bs$violence_yd_bs = ela_in_abcd_bs$ksads_ptsd_raw_766_p


#mental illness
#Life Events Family member had mental/emotional problem? ple_mh_y
ela_in_abcd_bs$mi_yd_bs = ela_in_abcd_bs$ple_mh_y_bs

#Incarcerated Relative
#Life Events Someone in the family was arrested? ple_arrest_y_bs
#Life Events Parents/caregiver got into trouble with the law? ple_law_y_bs
#Life Events One of the parents/caregivers went to jail? ple_jail_y_bs

incar_sum_y_bs_data = c("ple_arrest_y_bs", "ple_law_y_bs", "ple_jail_y_bs")

ela_in_abcd_bs$incar_sum_y = apply(ela_in_abcd_bs[incar_sum_y_bs_data],1, sum_score)

ela_in_abcd_bs$incar_yd_bs = ifelse(ela_in_abcd_bs$incar_sum_y == 0, 0, 1)

#Substance Use
#Life Events Family member had drug and/or alcohol problem?
ela_in_abcd_bs$su_yd_bs=ela_in_abcd_bs$ple_sud_y_bs

#*SUM SCORE*
ela_in_abcd_bs$aces_youth_bs = ela_in_abcd_bs$abuse_phy_yd_bs + ela_in_abcd_bs$abuse_sex_yd_bs + ela_in_abcd_bs$abuse_emo_yd_bs + ela_in_abcd_bs$neglect_phy_yd_bs + ela_in_abcd_bs$neglect_emo_yd_bs + ela_in_abcd_bs$divorce_yd_bs + ela_in_abcd_bs$violence_yd_bs + ela_in_abcd_bs$mi_yd_bs + ela_in_abcd_bs$incar_yd_bs + ela_in_abcd_bs$su_yd_bs

```

```{r yr1-yd}
#yr 1 

#abuse_phy
#only parent report available
#cannot be updated at Year 1 due to no KSADS

#abuse_sex
#only parent report available
#cannot be updated at Year 1 due to no KSADS

#abuse_emo
#Family Environment Family members often criticize each other. fes_youth_q5
#utilizing youth only to ensure youth exposure

ela_in_abcd_yr1$abuse_emo_yd_yr1 = ela_in_abcd_yr1$fes_youth_q5 

#neglect_phy
#only parent report available
#Demographics Needed food but couldn't afford to buy it or couldn't afford to go out to get it?P demo_fam_exp1_v2_l/demo_fam_exp1_v2
#Demographics Had someone who needed to see a doctor or go to the hospital but didn't go because you could not afford it?P demo_fam_exp6_v2/demo_fam_exp6_v2_l
#Demographics Had someone who needed a dentist but couldn't go because you could not afford it?P demo_fam_exp7_v2/demo_fam_exp7_v2_l

#create dataframe with physical neglect items
neglect_phy_sum_yr1_data = c("demo_fam_exp1_v2_l", "demo_fam_exp6_v2_l", "demo_fam_exp7_v2_l")

#sum the items as long as they're not all NA
ela_in_abcd_yr1$neglect_phy_sum = apply(ela_in_abcd_yr1[neglect_phy_sum_yr1_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_yr1$neglect_phy_yd_yr1 = ifelse(ela_in_abcd_yr1$neglect_phy_sum == 0, 0, 1)

#neglect_emo
#CRPBI parent *CRPBI Mean Parent (primary caregiver)Y crpbi_y_ss_parent_ace
ela_in_abcd_yr1$neglect_emo_yd_yr1 = ela_in_abcd_yr1$crpbi_y_ss_parent_ace

#divorce
#Life Events Parents separated or divorced?
ela_in_abcd_yr1$divorce_yd_yr1 = ela_in_abcd_yr1$ple_separ_y

#violence in home (violence)
#only parent report available
#cannot be updated at Year 1 due to no KSADS

#mental illness (mi)
#Life Events Family member had mental/emotional problem? ple_mh_y
ela_in_abcd_yr1$mi_yd_yr1 = ela_in_abcd_yr1$ple_mh_y

#Incarcerated Relative
#Life Events Someone in the family was arrested? ple_arrest_y_yr1
#Life Events Parents/caregiver got into trouble with the law? ple_law_y_yr1
#Life Events One of the parents/caregivers went to jail? ple_jail_y_yr1

#create dataframe with incarceration items
incar_sum_y_yr1_data = c("ple_arrest_y" , "ple_law_y"  , "ple_jail_y")

#sum the items as long as they're not all NA
ela_in_abcd_yr1$incar_sum_y = apply(ela_in_abcd_yr1[incar_sum_y_yr1_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_yr1$incar_yd_yr1 = ifelse(ela_in_abcd_yr1$incar_sum_y == 0, 0, 1)

#Substance Use
#Life Events Family member had drug and/or alcohol problem?
ela_in_abcd_yr1$su_yd_yr1 = ela_in_abcd_yr1$ple_sud_y

```

```{r yr2-yd}
#yr 2 

#abuse_phy
#only parent report available
#KSADS	Shot, stabbed, or beaten brutally by a non-family member	ksads_ptsd_raw_761_p
#KSADS	Shot, stabbed, or beaten brutally by a grown up in the home	ksads_ptsd_raw_762_p
#KSADS	Beaten to the point of having bruises by a grown up in the home	ksads_ptsd_raw_763_p
#KSADS	A non-family member threatened to kill your child	ksads_ptsd_raw_764_p
#KSADS	A family member threatened to kill your child	ksads_ptsd_raw_765_p

#create dataframe with physical abuse items
abuse_phy_yr2_data = c("ksads_ptsd_raw_761_p"  , "ksads_ptsd_raw_762_p", "ksads_ptsd_raw_763_p", "ksads_ptsd_raw_764_p", "ksads_ptsd_raw_765_p")

#sum the items as long as they're not all NA
ela_in_abcd_yr2$abuse_phy_sum = apply(ela_in_abcd_yr2[abuse_phy_yr2_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_yr2$abuse_phy_yd_yr2 = ifelse(ela_in_abcd_yr2$abuse_phy_sum == 0, 0, 1)

#abuse_sex
#only parent report available
#KSADS A grown up in the home touched your child in their privates, had your child touch their privates, or did other sexual things to your childP ksads_ptsd_raw_767_p
#KSADS An adult outside your family touched your child in their privates, had your child touch their privates or did other sexual things to your childP ksads_ptsd_raw_768_p
#KSADS A peer forced your child to do something sexuallyP ksads_ptsd_raw_769_p

#create dataframe with sexual abuse items
abuse_sex_yr2_data = c("ksads_ptsd_raw_767_p"  , "ksads_ptsd_raw_768_p", "ksads_ptsd_raw_769_p")

#sum the items as long as they're not all NA
ela_in_abcd_yr2$abuse_sex_sum = apply(ela_in_abcd_yr2[abuse_sex_yr2_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_yr2$abuse_sex_yd_yr2 = ifelse(ela_in_abcd_yr2$abuse_sex_sum == 0, 0, 1)

#abuse_emo
#Family Environment Family members often criticize each other. fes_youth_q5
#no recode necessary (binary yes/no variable)
ela_in_abcd_yr2$abuse_emo_yd_yr2 = ela_in_abcd_yr2$fes_youth_q5 

#neglect_phy
#only parent report available
#Demographics Needed food but couldn't afford to buy it or couldn't afford to go out to get it?P demo_fam_exp1_v2_l/demo_fam_exp1_v2
#Demographics Had someone who needed to see a doctor or go to the hospital but didn't go because you could not afford it?P demo_fam_exp6_v2/demo_fam_exp6_v2_l
#Demographics Had someone who needed a dentist but couldn't go because you could not afford it?P demo_fam_exp7_v2/demo_fam_exp7_v2_l

#create dataframe with physical neglect items
neglect_phy_yr2_data = c("demo_fam_exp1_v2_l"  , "demo_fam_exp6_v2_l", "demo_fam_exp7_v2_l")

#sum the items as long as they're not all NA
ela_in_abcd_yr2$neglect_phy_sum = apply(ela_in_abcd_yr2[neglect_phy_yr2_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_yr2$neglect_phy_yd_yr2 = ifelse(ela_in_abcd_yr2$neglect_phy_sum == 0, 0, 1)

#neglect_emo
#CRPBI parent *CRPBI Mean Parent (primary caregiver)Y crpbi_y_ss_parent_ace
#only ~5000 with scores, so without further information, omitting for year 2

#divorce
#Life Events Parents separated or divorced?
ela_in_abcd_yr2$divorce_yd_yr2 = ela_in_abcd_yr2$ple_separ_y

#violence in home (violence)
#only parent report available
#KSADS Witness the grownups in the home push, shove or hit one another 	ksads_ptsd_raw_766_p
ela_in_abcd_yr2$violence_yd_yr2 = ela_in_abcd_yr2$ksads_ptsd_raw_766_p

#mental illness (mi)
#Life Events Family member had mental/emotional problem?
ela_in_abcd_yr2$mi_yd_yr2 = ela_in_abcd_yr2$ple_mh_y

#Incarcerated Relative
#Life Events Someone in the family was arrested?
#Life Events Parents/caregiver got into trouble with the law?
#Life Events One of the parents/caregivers went to jail?

#create dataframe with incarceration items
incar_yr2_y_data = c("ple_arrest_y", "ple_law_y", "ple_jail_y")

#sum the items as long as they're not all NA
ela_in_abcd_yr2$incar_sum_y = apply(ela_in_abcd_yr2[incar_yr2_y_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_yr2$incar_yd_yr2 = ifelse(ela_in_abcd_yr2$incar_sum_y == 0, 0, 1)

#Substance Use
#Life Events Family member had drug and/or alcohol problem?
ela_in_abcd_yr2$su_yd_yr2 = ela_in_abcd_yr2$ple_sud_y

```

```{r y3-yd}
#yr 3 
#only parent report available
#not enough ksads data yr3

#abuse_sex
#only parent report available
#not enough ksads data yr3

#abuse_emo
#Family Environment Family members often criticize each other. fes_youth_q5
#utilizing youth only to ensure youth exposure
ela_in_abcd_yr3$abuse_emo_yd_yr3 = ela_in_abcd_yr3$fes_youth_q5 

#neglect_phy
#only parent report available
#Demographics Needed food but couldn't afford to buy it or couldn't afford to go out to get it?P demo_fam_exp1_v2_l/demo_fam_exp1_v2
#Demographics Had someone who needed to see a doctor or go to the hospital but didn't go because you could not afford it?P demo_fam_exp6_v2/demo_fam_exp6_v2_l
#Demographics Had someone who needed a dentist but couldn't go because you could not afford it?P demo_fam_exp7_v2/demo_fam_exp7_v2_l

#create dataframe with phys neglect items
neglect_phy_yr3_data = c("demo_fam_exp1_v2_l" , "demo_fam_exp6_v2_l", "demo_fam_exp7_v2_l")

#sum the items as long as they're not all NA
ela_in_abcd_yr3$neglect_phy_sum = apply(ela_in_abcd_yr3[neglect_phy_yr3_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_yr3$neglect_phy_yd_yr3 = ifelse(ela_in_abcd_yr3$neglect_phy_sum == 0, 0, 1)

#neglect_emo
#CRPBI parent *CRPBI Mean Parent (primary caregiver)Y crpbi_y_ss_parent_ace
ela_in_abcd_yr3$neglect_emo_yd_yr3 = ela_in_abcd_yr3$crpbi_y_ss_parent_ace

#divorce
#Life Events Parents separated or divorced?
ela_in_abcd_yr3$divorce_yd_yr3 = ela_in_abcd_yr3$ple_separ_y

#violence in home (violence)
#only parent report available
#not enough ksads data for yr3

#mental illness (mi)
#Life Events Family member had mental/emotional problem?
ela_in_abcd_yr3$mi_yd_yr3 = ela_in_abcd_yr3$ple_mh_y

#Incarcerated Relative
#Life Events Someone in the family was arrested? ple_arrest_y_yr3
#Life Events Parents/caregiver got into trouble with the law? ple_law_y_yr3
#Life Events One of the parents/caregivers went to jail? ple_jail_y_yr3

#create dataframe with incarceration items 
incar_yr3_y_data = c("ple_arrest_y", "ple_law_y", "ple_jail_y")

#sum the items as long as they're not all NA 
ela_in_abcd_yr3$incar_sum_y = apply(ela_in_abcd_yr3[incar_yr3_y_data],1, sum_score)

#binary recode as anything > 0 is 1
ela_in_abcd_yr3$incar_yd_yr3 = ifelse(ela_in_abcd_yr3$incar_sum_y == 0, 0, 1)

#Substance Use
#Life Events Family member had drug and/or alcohol problem?
ela_in_abcd_yr3$su_yd_yr3 = ela_in_abcd_yr3$ple_sud_y

```

```{r}
##LONGITUDINAL SCORE CREATION

#join annual follow-up scores to Baseline
#pulling just scores, no eventnames (wide format)
bs_scores = ela_in_abcd_bs[c("src_subject_id", "aces_youth_bs", "abuse_phy_yd_bs", "abuse_sex_yd_bs", "abuse_emo_yd_bs", "neglect_phy_yd_bs", "neglect_emo_yd_bs", "divorce_yd_bs", "violence_yd_bs", "mi_yd_bs", "incar_yd_bs", "su_yd_bs")]                  
yr1_scores = ela_in_abcd_yr1[c("src_subject_id", "abuse_emo_yd_yr1", "neglect_phy_yd_yr1", "neglect_emo_yd_yr1", "divorce_yd_yr1", "mi_yd_yr1", "incar_yd_yr1", "su_yd_yr1")]
yr2_scores = ela_in_abcd_yr2[c("src_subject_id", "abuse_phy_yd_yr2", "abuse_sex_yd_yr2", "abuse_emo_yd_yr2", "neglect_phy_yd_yr2", "divorce_yd_yr2", "violence_yd_yr2", "mi_yd_yr2", "incar_yd_yr2", "su_yd_yr2")]     
yr3_scores = ela_in_abcd_yr3[c("src_subject_id", "abuse_emo_yd_yr3", "neglect_phy_yd_yr3", "neglect_emo_yd_yr3", "divorce_yd_yr3", "mi_yd_yr3", "incar_yd_yr3", "su_yd_yr3")]

yr1 = left_join(bs_scores,yr1_scores, by=c("src_subject_id")) 
yr2 = left_join(yr1, yr2_scores, by=c("src_subject_id")) 
ela_final = left_join(yr2, yr3_scores, by=c("src_subject_id")) 

#save bs data
#prep for return to long format
bs_scores$eventname="baseline_year_1_arm_1"
lookup = c(aces_proxy_yd="aces_youth_bs", abuse_phy = "abuse_phy_yd_bs", abuse_sex = "abuse_sex_yd_bs", abuse_emo = "abuse_emo_yd_bs", neglect_phy = "neglect_phy_yd_bs", neglect_emo = "neglect_emo_yd_bs", divorce = "divorce_yd_bs", violence = "violence_yd_bs", mental = "mi_yd_bs", incar = "incar_yd_bs", sud = "su_yd_bs")
bs_scores=rename(bs_scores, all_of(lookup))

#building score - adding items by timepoint
#for each domain, if current and any/all previous timepoints 0, will be 0; otherwise, 1

#yr1 by domain
ela_final$abuse_phy_yr1_build = ifelse(ela_final$abuse_phy_yd_bs == 0, 0, 1)
ela_final$abuse_sex_yr1_build = ifelse(ela_final$abuse_sex_yd_bs == 0, 0, 1)
ela_final$abuse_emo_yr1_build = ifelse(ela_final$abuse_emo_yd_bs == 0 & ela_final$abuse_emo_yd_yr1 == 0, 0, 1)
ela_final$neglect_phy_yr1_build = ifelse(ela_final$neglect_phy_yd_bs == 0 & ela_final$neglect_phy_yd_yr1 == 0, 0, 1)
ela_final$neglect_emo_yr1_build = ifelse(ela_final$neglect_emo_yd_bs == 0 & ela_final$neglect_emo_yd_yr1 == 0, 0, 1)
ela_final$divorce_yr1_build = ifelse(ela_final$divorce_yd_bs == 0 & ela_final$divorce_yd_yr1 == 0, 0, 1)
ela_final$violence_yr1_build = ifelse(ela_final$violence_yd_bs == 0, 0, 1)
ela_final$mi_yr1_build = ifelse(ela_final$mi_yd_bs == 0 & ela_final$mi_yd_yr1 == 0, 0, 1)
ela_final$incar_yr1_build = ifelse(ela_final$incar_yd_bs == 0 & ela_final$incar_yd_yr1 == 0, 0, 1)
ela_final$su_yr1_build = ifelse(ela_final$su_yd_bs == 0 & ela_final$su_yd_yr1 == 0, 0, 1)

#yr1 sum score
ela_final$aces_yr1_yd_build = ela_final$abuse_phy_yr1_build + ela_final$abuse_sex_yr1_build + ela_final$abuse_emo_yr1_build + ela_final$neglect_phy_yr1_build + ela_final$neglect_emo_yr1_build + ela_final$divorce_yr1_build + ela_final$violence_yr1_build + ela_final$mi_yr1_build + ela_final$incar_yr1_build + ela_final$su_yr1_build

#save yr1 data
yr1_scores_final = ela_final[c("src_subject_id", "aces_yr1_yd_build", "abuse_phy_yr1_build", "abuse_sex_yr1_build", "abuse_emo_yr1_build", "neglect_phy_yr1_build", "neglect_emo_yr1_build", "divorce_yr1_build", "violence_yr1_build", "mi_yr1_build", "incar_yr1_build", "su_yr1_build")]

#prep for return to long format
yr1_scores_final$eventname="1_year_follow_up_y_arm_1"
lookup = c(aces_proxy_yd = "aces_yr1_yd_build", abuse_phy = "abuse_phy_yr1_build", abuse_sex = "abuse_sex_yr1_build", abuse_emo = "abuse_emo_yr1_build", neglect_phy = "neglect_phy_yr1_build", neglect_emo = "neglect_emo_yr1_build", divorce = "divorce_yr1_build", violence = "violence_yr1_build", mental = "mi_yr1_build", incar = "incar_yr1_build", sud = "su_yr1_build")
yr1_scores_final=rename(yr1_scores_final, all_of(lookup))

#yr2 by domain
ela_final$abuse_phy_yr2_build = ifelse(ela_final$abuse_phy_yd_bs == 0 & ela_final$abuse_phy_yd_yr2 == 0, 0, 1)
ela_final$abuse_sex_yr2_build = ifelse(ela_final$abuse_sex_yd_bs == 0 & ela_final$abuse_sex_yd_yr2 == 0, 0, 1)
ela_final$abuse_emo_yr2_build = ifelse(ela_final$abuse_emo_yd_bs == 0 & ela_final$abuse_emo_yd_yr1 == 0 & ela_final$abuse_emo_yd_yr2 == 0, 0, 1)
ela_final$neglect_phy_yr2_build = ifelse(ela_final$neglect_phy_yd_bs == 0 & ela_final$neglect_phy_yd_yr1 == 0 & ela_final$neglect_phy_yd_yr2 == 0, 0, 1)
ela_final$neglect_emo_yr2_build = ifelse(ela_final$neglect_emo_yd_bs == 0 & ela_final$neglect_emo_yd_yr1 == 0, 0, 1)
ela_final$divorce_yr2_build = ifelse(ela_final$divorce_yd_bs == 0 & ela_final$divorce_yd_yr1 == 0 & ela_final$divorce_yd_yr2 == 0, 0, 1)
ela_final$violence_yr2_build = ifelse(ela_final$violence_yd_bs == 0 & ela_final$violence_yd_yr2 == 0, 0, 1)
ela_final$mi_yr2_build = ifelse(ela_final$mi_yd_bs == 0 & ela_final$mi_yd_yr1 == 0 & ela_final$mi_yd_yr2 == 0, 0, 1)
ela_final$incar_yr2_build = ifelse(ela_final$incar_yd_bs == 0 & ela_final$incar_yd_yr1 == 0 & ela_final$incar_yd_yr2 == 0, 0, 1)
ela_final$su_yr2_build = ifelse(ela_final$su_yd_bs == 0 & ela_final$su_yd_yr1 == 0 & ela_final$su_yd_yr2 == 0, 0, 1)

#yr2 sum score
ela_final$aces_yr2_yd_build = ela_final$abuse_phy_yr2_build + ela_final$abuse_sex_yr2_build + ela_final$abuse_emo_yr2_build + ela_final$neglect_phy_yr2_build + ela_final$neglect_emo_yr2_build + ela_final$divorce_yr2_build + ela_final$violence_yr2_build + ela_final$mi_yr2_build + ela_final$incar_yr2_build + ela_final$su_yr2_build

#save yr2 data
yr2_scores_final = ela_final[c("src_subject_id", "aces_yr2_yd_build", "abuse_phy_yr2_build", "abuse_sex_yr2_build", "abuse_emo_yr2_build", "neglect_phy_yr2_build", "neglect_emo_yr2_build", "divorce_yr2_build", "violence_yr2_build", "mi_yr2_build", "incar_yr2_build", "su_yr2_build")]

#prep for return to long format
yr2_scores_final$eventname="2_year_follow_up_y_arm_1"
lookup = c(aces_proxy_yd = "aces_yr2_yd_build", abuse_phy = "abuse_phy_yr2_build", abuse_sex = "abuse_sex_yr2_build", abuse_emo = "abuse_emo_yr2_build", neglect_phy = "neglect_phy_yr2_build", neglect_emo = "neglect_emo_yr2_build", divorce = "divorce_yr2_build", violence = "violence_yr2_build", mental = "mi_yr2_build", incar = "incar_yr2_build", sud = "su_yr2_build")
yr2_scores_final=rename(yr2_scores_final, all_of(lookup))

#yr3 by domain
ela_final$abuse_phy_yr3_build = ifelse(ela_final$abuse_phy_yd_bs == 0 & ela_final$abuse_phy_yd_yr2 == 0, 0, 1)
ela_final$abuse_sex_yr3_build = ifelse(ela_final$abuse_sex_yd_bs == 0 & ela_final$abuse_sex_yd_yr2 == 0, 0, 1)
ela_final$abuse_emo_yr3_build = ifelse(ela_final$abuse_emo_yd_bs == 0 & ela_final$abuse_emo_yd_yr1 == 0 & ela_final$abuse_emo_yd_yr2 == 0 & ela_final$abuse_emo_yd_yr3 == 0, 0, 1)
ela_final$neglect_phy_yr3_build = ifelse(ela_final$neglect_phy_yd_bs == 0 & ela_final$neglect_phy_yd_yr1 == 0 & ela_final$neglect_phy_yd_yr2 == 0 & ela_final$neglect_phy_yd_yr3 == 0, 0, 1)
ela_final$neglect_emo_yr3_build = ifelse(ela_final$neglect_emo_yd_bs == 0 & ela_final$neglect_emo_yd_yr1 == 0 & ela_final$neglect_emo_yd_yr3 == 0, 0, 1)
ela_final$divorce_yr3_build = ifelse(ela_final$divorce_yd_bs == 0 & ela_final$divorce_yd_yr1 == 0 & ela_final$divorce_yd_yr2 == 0 & ela_final$divorce_yd_yr3 == 0, 0, 1)
ela_final$mi_yr3_build = ifelse(ela_final$mi_yd_bs == 0 & ela_final$mi_yd_yr1 == 0 & ela_final$mi_yd_yr2 == 0 & ela_final$mi_yd_yr3 == 0, 0, 1)
ela_final$incar_yr3_build = ifelse(ela_final$incar_yd_bs == 0 & ela_final$incar_yd_yr1 == 0 & ela_final$incar_yd_yr2 == 0 & ela_final$incar_yd_yr3 == 0, 0, 1)
ela_final$su_yr3_build = ifelse(ela_final$su_yd_bs == 0 & ela_final$su_yd_yr1 == 0 & ela_final$su_yd_yr2 == 0 & ela_final$su_yd_yr3 == 0, 0, 1)
ela_final$violence_yr3_build = ifelse(ela_final$violence_yd_bs == 0 & ela_final$violence_yd_yr2 == 0, 0, 1)

#yr3 sum score
ela_final$aces_yr3_yd_build = ela_final$abuse_phy_yr3_build + ela_final$abuse_sex_yr3_build + ela_final$abuse_emo_yr3_build + ela_final$neglect_phy_yr3_build + ela_final$neglect_emo_yr3_build + ela_final$divorce_yr3_build + ela_final$violence_yr3_build + ela_final$mi_yr3_build + ela_final$incar_yr3_build + ela_final$su_yr3_build

#save yr3 data
yr3_scores_final = ela_final[c("src_subject_id", "aces_yr3_yd_build", "abuse_phy_yr3_build", "abuse_sex_yr3_build", "abuse_emo_yr3_build", "neglect_phy_yr3_build", "neglect_emo_yr3_build", "divorce_yr3_build", "violence_yr3_build", "mi_yr3_build", "incar_yr3_build", "su_yr3_build")]

#prep for return to long format
yr3_scores_final$eventname="3_year_follow_up_y_arm_1"
lookup = c(aces_proxy_yd = "aces_yr3_yd_build", abuse_phy = "abuse_phy_yr3_build", abuse_sex = "abuse_sex_yr3_build", abuse_emo = "abuse_emo_yr3_build", neglect_phy = "neglect_phy_yr3_build", neglect_emo = "neglect_emo_yr3_build", divorce = "divorce_yr3_build", violence = "violence_yr3_build", mental = "mi_yr3_build", incar = "incar_yr3_build", sud = "su_yr3_build")
yr3_scores_final=rename(yr3_scores_final, all_of(lookup))

#combine datasets
step1=rbind(bs_scores,yr1_scores_final)
step2=rbind(step1,yr2_scores_final)
ela_in_abcd_scores=rbind(step2,yr3_scores_final)

#save data
write.csv(ela_in_abcd_scores, "aces_proxy_youth_abcd.csv")
```

